---
title: "Prep_TrackingDataWrangling"
author: "Pauliina Vuorinen"
date: "18/03/2022-"
output: html_document
library: "S1_code_library.bib"
---

## Introduction

The purpose of this script is to clean the raw reading behaviour tracking data for our analysis.

## Setup

Load libraries, set working directory

```{r, setup}
library(psych)
library(tidyverse)
library(tidyr)
library(dplyr)
getwd() # check that working directory is ~/Short_Story_Reading_Behaviour_Public
```

## Load data and check variable types

Load raw data from Data -folder and change variable types. See documents in Info -folder for information on different tracking events.

```{r, load data}
raw_tracking_data <-
    read.csv(
        "Data/preprocessed_raw_tracking_data.csv",
        header = TRUE,
        sep = ";",
        dec = ","
    )
head(raw_tracking_data)
```

```{r, change variable types}
str(raw_tracking_data)
names(raw_tracking_data)[names(raw_tracking_data) == "BookId"] <- "StoryId"
source("Functions/Functions_VariableTypeConversion.R")

## turn user and story indicators into factors
raw_tracking_data[, c(
    "UserId",
    "StoryId"
)] <- convert.magic(
    raw_tracking_data[, c(
        "UserId",
        "StoryId"
    )],
    "factor"
)
## turn reading block (page view) number  into an ordered variable

raw_tracking_data$ReadingBlockNumber <-
    as.ordered(
        raw_tracking_data$ReadingBlockNumber
    )
## fix date and time variable types
raw_tracking_data$Date <-
    as.Date(
        raw_tracking_data$Date,
        format = "%d.%m.%Y"
    )
raw_tracking_data$Time <-
    gsub(
        x = raw_tracking_data$Time, ",", "."
    )
op <-
    options(digits.secs = 3)
raw_tracking_data$Time <-
    strptime(
        raw_tracking_data$Time,
        format = "%H:%M:%OS"
    )
```

```{r, order df}
# order by User, date, time and event id
raw_tracking_data <-
    raw_tracking_data[
        with(
            raw_tracking_data,
            order(UserId, Date, Time, Id)
        ),
    ]
```

## Exclusion criteria

Sixty-two participants were recruited to take part in the study. Two of the participants did not finish the study, and so they are excluded from the sample. The final *n* = 60.

```{r, exclusion}
raw_tracking_data <-
    filter(
        raw_tracking_data,
        (UserId != "33") &
            (UserId != "61")
    )
```

## Missing and erroneous data values

Fix unrealistic and misleading data values in the tracking data. These errors arise from accidental miscoding during the collection of the tracking data, or alternative data formatting.

1. Change navigation block numbers (continuous navigation in one direction) recorded as zero into NA values. A navigation block number is only recorded as zero if the participant is not navigating the document. This is misleading, and so zeros are turned to NAs.
```{r, fix navigation block formatting}
raw_tracking_data[, "NavigationBlockNumber"][raw_tracking_data[, "NavigationBlockNumber"] == 0] <- NA
```

2. Similarly to navigation block numbers, reading block numbers (page views) are recorded as zero when the participant is not viewing a page. This is misleading, and so zeros are turned to NAs.
```{r, fix reading block formatting}
raw_tracking_data[, "ReadingBlockNumber"][raw_tracking_data[, "ReadingBlockNumber"] == 0] <- NA
```

3. Inactive timeout events are erroneously coded as 'engagements' in the raw data. The inactive timeout indicates that the page view has been masked due to > 5mins of inactivity, and therefore, these event types should be coded as 'disengagements'.
```{r, fix inactiveTimeout engagement type}
raw_tracking_data[raw_tracking_data$Type == "inactiveTimeout", ]$Engagement <- "Disengagement"
```

### Adjustments to event and navigation block durations

Event durations are important not only because they can tell us about how long an event lasted, but 'DurationMinutes' is also used to create cumulative durations (such as 'EngagementDuration'). The event durations are calculated by deducting an event's start time from the following event's start time. Therefore, the durations of events in the middle and the beginning of a reading session are accurate as long as all events are present in the tracking data. The last event in a reading session, however, has an excessive duration because duration is miscalculated by comparing the event to the next one that is in a new reading session. This excessive duration is adjusted in two different ways, depending on whether the event is categorised as a 'disengagement' or an 'engagement'.

```{r, find last event in reading session}
# an event is the last one in a reading session if:
## the current session != next, or the current user != next
raw_tracking_data$IsLastEventInReadingSession <-
    (
        (raw_tracking_data$ReadingSessionNumber !=
            lead(raw_tracking_data$ReadingSessionNumber, 1)) |
            (raw_tracking_data$UserId !=
                lead(raw_tracking_data$UserId, 1))
    )
# last event in df assigned manually
raw_tracking_data$IsLastEventInReadingSession <-
    replace_na(raw_tracking_data$IsLastEventInReadingSession, TRUE)
```

124 of the last reading session events are disengagements and five are engagements.

```{r, check last event in reading session engagement types}
table(
    raw_tracking_data$Engagement,
    raw_tracking_data$IsLastEventInReadingSession
)
```

**Disengagements.** If the last event in a reading session is a disengagement, such as 'close' or 'logout' (see event types in info folder, document X), the duration should be recorded as 0. A disengagement at the end of a reading session indicates of the participants' intention to leave the e-reader system, and so it's duration is arbitrary.

**Engagements.** Engagements that have a duration of >5min indicate of a missing tracking event. A reading session ending in an 'engagement' indicates that the final disengagement event (triggered by the closure of the e-reader system, for example, event of type 'logout') is missing. Similarly, an engagement event that has a duration of >5min in the middle of a reading session is indicative of a missing event: the e-reader system automatically masks the text after 5 minutes of inactivity (event type 'blur'), if an engagement event lasts for longer than 5 minutes, the text mask event may have not been recorded in the tracking data.

Missing events occur when the system doesn't have the time to record an event, for example when the e-reader system is abruptly closed or when network connection fails (see Note). Indeed, correctly recorded engagements should last for a maximum of 5 minutes before the e-reader system masks the text. Therefore, excessive engagement durations are checked by filtering the dataset for engagements that last for longer than 5.1 minutes.

*Note: the e-reader system works on the client side, but no temporary data storage is implemented to store events collected during network connection issues and so these events are lost in case of network connection failures.*

```{r, find excessive engagement durations}
excessive_engagement_durations_subdata <-
    filter(
        raw_tracking_data,
        (Engagement == "Engagement") &
            (DurationMinutes > 5.1)
    )
table(excessive_engagement_durations_subdata$IsLastEventInReadingSession) # CHECK
```

The excessive durations are adjusted by creating new variables:
1. 'AdjustedEventDuration' shows the suggested new event duration
    * Set to 5 minutes for engagement events and to 0 for non-engagement events.

```{r, adjust to event durations}
raw_tracking_data$AdjustedEventDuration <-
    ifelse(
        ((raw_tracking_data$Engagement == "Engagement") &
            (raw_tracking_data$DurationMinutes > 5.1)),
        5,
        ifelse(
            ((raw_tracking_data$IsLastEventInReadingSession &
                raw_tracking_data$Engagement != "Engagement")),
            0
        )
    )
```

2. 'AdjustedNavigationBlockDuration' shows the suggested adjustment to NavigationBlockDuration
    * Excessive engagement events' duration that exceeds 5 minutes is deducted from the navigation block duration, whereas excessive non-engagement events' entire duration is deducted from the navigation block duration.

```{r, adjust navigation block durations}
raw_tracking_data$AdjustedNavigationBlockDuration <-
    ifelse(
        ((raw_tracking_data$AdjustedEventDuration != NA) &
            (raw_tracking_data$Engagement == "Engagement")),
        (raw_tracking_data$NavigationBlockDuration -
            (raw_tracking_data$DurationMinutes - 5)),
        ifelse(
            ((raw_tracking_data$AdjustedEventDuration != NA) &
                (raw_tracking_data$Engagement != "Engagement")),
            (raw_tracking_data$NavigationBlockDuration -
                raw_tracking_data$DurationMinutes)
        )
    )
```

### Identification of inaccurately categorised disengagements

Events were given engagement labels in the data pre-processing stage (see file __ in the Info folder for detail), however, some rare events were mistakenly classified as disengagements. During the opening of the e-reader system, an event was occasionally categorised as a 'disengagement' even though the text was visible to the participants. These events are first identified by inspection of 'unusual disengagements' that don't have a clear disengagement-linked event type (see information on events in ___ in the Info folder). Ten events were identified as inaccurately categorised and so these events are corrected into engagements.

```{r, inaccurately categorised disengagements}
raw_tracking_data$IsUnusualDisengagement <-
    ((raw_tracking_data$Engagement == "Disengagement") &
        (raw_tracking_data$Type != "blur" &
            raw_tracking_data$Type != "openMenu" &
            raw_tracking_data$Type != "inactiveTimeout" &
            raw_tracking_data$Type != "close" &
            raw_tracking_data$Type != "logout" &
            raw_tracking_data$Type != "openBook" &
            raw_tracking_data$Type != "bookDialogOpen" &
            raw_tracking_data$Type != "openInformation"))
raw_tracking_data$Engagement <-
    ifelse(data$IsUnusualDisengagement &
        raw_tracking_data$IsOpeningEvent &
        raw_tracking_data$IsPageOpen &
        !raw_tracking_data$IsBlurred &
        !raw_tracking_data$IsDialogOpen &
        !raw_tracking_data$IsMenuOpen &
        !raw_tracking_data$IsInactive,
    "Engagement",
    raw_tracking_data$Engagement
    )
```

### Reading session artefacts

The data analysis for the current study was focused on engaged reading, and so very short reading sessions or sessions that largely consisted of disengagement events only were not considered to be meaningful. These reading sessions were likely to occur if the participant attempts to resume a reading session and becomes quickly distracted by other things, or if they accidentally open the e-reader.

We expected that an intentional reading session would last more than a minute, and it would consist of less than 95% of disengagement. To ensure that this threshold was meaningful and did not remove reading sessions of interest, we identified potential reading session artefacts with these thresholds and compared the artefacts identified to visualisations of each participants' reading sessions (see timeline figures in folder Figu). The comparison showed that these two thresholds could identify most of the reading sessions that seemed to consist of little reading. Additional testing showed that some of the remaining reading sessions mostly consisted of a combination of disengagements and 'dialog' -events (dialog events refer to the opening of the menu), and so another threshold was added: a reading session of interest should consist of less than 80% of disengagement and dialog. Comparison to the figures indicated that these three thresholds were sufficient in detecting reading session artefacts.

```{r, identify reading session artefacts}
# prepare a new df
UserList <- levels(raw_tracking_data$UserId)
SessionDetails <- data.frame(
    UserId = factor(),
    ReadingSessionNumber = numeric(),
    SessionLength = numeric(),
    DisengagementDuration = numeric(),
    DialogDuration = numeric(),
    SlowReadingDuration = numeric(),
    EngagementProportion = numeric(),
    DisengagementProportion = numeric(),
    DialogAndDisengagementProportion = numeric()
)
firstrow <- 1
raw_tracking_data$ReadingSessionNumber <-
    as.numeric(
        as.character(
            raw_tracking_data$ReadingSessionNumber
        )
    )

# calculate session details to the new df
for (user in UserList) {
    subset_df1 <- filter(raw_tracking_data, UserId == user)
    for (session in unique(subset_df1$ReadingSessionNumber)) {
        subset_df2 <- filter(subset_df1, ReadingSessionNumber == session)
        temporary_df <- subset_df2 %>%
            summarise(
                SessionLength =
                    sum(AdjustedEventDuration, na.rm = TRUE),
                DisengagementDuration =
                    sum(
                        subset_df2[subset_df2$Engagement == "Disengagement", ]$AdjustedEventDuration,
                        na.rm = TRUE
                    ),
                DialogDuration =
                    sum(
                        subset_df2[subset_df2$Engagement == "Dialog", ]$AdjustedEventDuration,
                        na.rm = TRUE
                    ),
                SlowReadingDuration =
                    sum(
                        subset_df2[subset_df2$NavBlockSpeedLabel == "SlowReading", ]$AdjustedEventDuration,
                        na.rm = TRUE
                    ),
                EngagementProportion =
                    (
                        sum(
                            subset_df2[subset_df2$Engagement == "Engagement", ]$AdjustedEventDuration,
                            na.rm = TRUE
                        )
                        / SessionLength
                    ),
                DisengagementProportion =
                    (DisengagementDuration / SessionLength),
                DialogAndDisengagementProportion =
                    ((DisengagementDuration + DialogDuration) / SessionLength)
            )
        temporary_df$UserId <- user
        temporary_df$ReadingSessionNumber <- session
        rows <- nrow(temporary_df)
        SessionDetails <- bind_rows(SessionDetails, temporary_df)
        firstrow <- firstrow + rows
    }
}

# identify potential reading session artefacts based on thresholds
SessionDetails$PotentialReadingSessionArtefact <-
    ifelse(
        ((SessionDetails$SessionLength < 1) |
            (SessionDetails$DisengagementProportion > 0.95) |
            (SessionDetails$DialogAndDisengagementProportion > 0.8)),
        TRUE,
        FALSE
    )
Artefacts <-
    filter(
        SessionDetails,
        PotentialReadingSessionArtefact
    )
```

A new core dataframe was created in which reading session artefacts were removed.

```{r, remove reading session artefacts}
tracking_data_artefacts_removed <-
    filter(
        raw_tracking_data,
        !PossibleArtefactSession
    )
# order the new df
tracking_data_artefacts_removed <-
    tracking_data_artefacts_removed[with(tracking_data_artefacts_removed, order(UserId, Date, Time, Id)), ]
```

## Cumulative time in a reading session

Add a variable in the dataframe that tells us the cumulative time in a reading session for each event.

```{r, add cumulative time}
tracking_data_artefacts_removed <-
    tracking_data_artefacts_removed %>%
    group_by(UserId, ReadingSessionNumber) %>%
    mutate(
        CumReadingSessionTime = cumsum(AdjustedEventDuration)
    )
```


## Save wrangled tracking data

The wrangled tracking data is saved for usage in other scripts to compute reading behaviour measures and to eventually analyse the results.
The dataset has already been saved, and so the below r code chunk is not run.

```{r, eval=FALSE, echo=TRUE, save wrangled data}
write.csv2("wrangled_tracking_data.csv", header = TRUE, dec = ",", sep = ";")
```